---
title: "1.1 — The Tools of Microeconomics — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Dates
*Section 1: Monday, August 24, 2022*

*Section 2: Tuesday, August 25, 2022*
:::

::: callout-warning
# Upcoming Assignment
Please complete the [preliminary survey](/assignments/00-preliminary-survey.qmd) by next class.
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

<span class="newthought">Welcome</span> to ECON 306 — Microeconomic Analysis! Today's lesson will be an overview of the content and the assignments of the course. Please read and familiarize yourself with the [syllabus](/syllabus). 

This is not just a “syllabus day,” as we need to hit the ground running, but today will give you a “big picture” look at this semester, and how to use microeconomic analysis, by focusing on the two main tools/models: **optimization** and **equilibrium**. We will dive right into optimization models next class.

## <i class="fas fa-book-reader fa-lg"></i> Readings

Today is introductory, but you may look at Chapter 1 if you wish, to remind you about the use of models in economics. Otherwise, please heed this timeless message:

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">A message to students from the Doggfather himself, <a href="https://twitter.com/SnoopDogg?ref_src=twsrc%5Etfw">@SnoopDogg</a> <a href="https://t.co/wsSANYv8u6">pic.twitter.com/wsSANYv8u6</a></p>&mdash; Ryan Briggs (@ryancbriggs) <a href="https://twitter.com/ryancbriggs/status/1293543531496308736?ref_src=twsrc%5Etfw">August 12, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

Please note going forward, the lesson numbers and topics (e.g. 1.1) on this website are *my* design, and will not match up with the textbook!

### Optional/Referenced Readings

- <i class="fas fa-external-link-square-alt"></i> de Rohan, Anuschka, 2003, [“Why Dolphins are Deep Thinkers”](https://www.theguardian.com/science/2003/jul/03/research.science), *The Guardian*
- <i class="fas fa-external-link-square-alt"></i> Katzen et. al, 2021, [The nematode worm C. elegans chooses between bacterial foods exactly as if maximizing economic utility](https://www.biorxiv.org/content/10.1101/2021.04.25.441352v2), *BioRxiv*
- <i class="fas fa-external-link-square-alt"></i> Loch-Temzelides, Ted, 2021, ["Walrasian equilibrium behavior in nature"](https://www.pnas.org/doi/10.1073/pnas.2020961118), *PNAS* 118(27)
- <i class="fas fa-external-link-square-alt"></i> Serrano, Jody, 2021, [These Monkeys Understand Economics and Intentionally Steal High-Ticket Items to Barter for Better Food, Study Finds](https://gizmodo.com/these-monkeys-understand-economics-and-intentionally-st-1846078932?utm_medium=sharefromsite&utm_source=_facebook), *Gizmodo*

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.1-slides")
```

## <i class="fas fa-laptop-code"></i> Assignments

### Preliminary Math Survey Due Next Class

Please take the [preliminary survey](/assignment/00-preliminary-survey) on your math background by next class. This will help us all have a productive semester together.
