---
title: "1.10 — The Hecksher-Olin Model: Empirics — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Date

*Thursday, February 23, 2023*
:::

::: callout-warning
# Upcoming Assignment
[Problem Set 1](/assignment/01-problem-set) (1.1-1.8) is due by 11:59 PM Thursday February 23 on Blackboard Assignments.
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today, we cover the empirical findings and tests of the Hecksher-Ohlin model.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- Chapter 4.2 in Feenstra & Taylor

### Primary Sources Mentioned:

- [<i class="fas fa-file-pdf"></i> Leontief (1953), “Domestic Production and Foreign Trade; The American Capital Position Re-Examined”](/readings/Leontief-1953.pdf)
- [<i class="fas fa-file-pdf"></i> Bowen, Leamer, and Sveikauskas (1987), “Multicountry, Multifactor Tests of Factor Abundance Theory”](/readings/Bowen-1987.pdf)
- [<i class="fas fa-file-pdf"></i> Trefler (1995), “The Case of the Missing Trade and Other Mysteries”](/readings/Trefler-1995.pdf)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.10-slides")
```

## <i class="fas fa-laptop-code"></i> Assignments

### Problem Set 1 Due Thursday February 23

[Problem Set 1](/assignment/01-problem-set) (1.1-1.8) is due by 11:59 PM Thursday February 23 on Blackboard Assignments.
