---
title: "1.11 — New Trade Theory I — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Date

*Tuesday, February 28, 2023*
:::

::: callout-warning
# Upcoming Assignment
[Problem Set 1](/assignment/02-problem-set) (1.9-1.12) is due by 11:59 PM Thursday March 9 on Blackboard Assignments.
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today, we begin New Trade Theory.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- Chapter 6.4 in Feenstra & Taylor

We will cover 6.1-6.2 next class.

I also strongly suggest you consume Paul Krugman’s [2008 Economics Nobel Prize address](https://www.nobelprize.org/prizes/economic-sciences/2008/krugman/lecture/) in some format. There is a video, his slides, and his written lecture posted there. I will be drawing from some of that today.

# Primary Sources Mentioned:

- [<i class="fas fa-file-pdf"></i> Krugman (1980), “Scale Economies, Product Differentiation, and Pattern of Trade”](/files/readings/Krugman-1980.pdf)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.10-slides")
```

## <i class="fas fa-laptop-code"></i> Assignments

### Problem Set 1 Due Thursday February 23

[Problem Set 1](/assignment/01-problem-set) (1.1-1.8) is due by 11:59 PM Thursday February 23 on Blackboard Assignments.
