---
title: "1.2 — The Smithean View of Trade: Divison of Labor — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Date

*Thursday, January 26, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we look at Adam Smith's understanding of what is the cause of the wealth of nations: the division of labor.

## <i class="fas fa-book-reader fa-lg"></i> Readings

### Recommended Reading

- [<i class="fas fa-external-link-square-alt"></i> Smith (1776), *Wealth of Nations*, Book I, Chs. 1-3](https://www.econlib.org/library/Smith/smWN.html)
- [<i class="fas fa-external-link-square-alt"></i> Safner (2023), “It All Beings with the Division of Labor: Meditations on what should be the central insight of economics”](https://increasingreturns.substack.com/p/it-all-begins-with-the-division-of)
- [<i class="fas fa-external-link-square-alt"></i> Safner (2023), “Division of Labor vs. the Invisible Hand: Two divergent trajectories of economic thought from two of its foundations”](https://increasingreturns.substack.com/p/division-of-labor-vs-the-invisible)
- [<i class="fas fa-external-link-square-alt"></i> Munger (2007), "I'll Stick With These: Some Sharp Observations on the Division of Labor"](https://www.econlib.org/library/Columns/y2007/Mungerpins.html)
- [<i class="fas fa-file-pdf"></i> Bauer (2000), "From Subsistence to Exchange"](/readings/Bauer-2000-Ch1.pdf)
- [<i class="fab fa-youtube"></i> TED Talk: Thomas Thwaites: How I Built a Toaster From Scratch"](https://www.ted.com/talks/thomas_thwaites_how_i_built_a_toaster_from_scratch?language=en)
- [<i class="fab fa-youtube"></i> YouTube: How to Make a $1500 Sandwich in Only 6 Months"](https://www.youtube.com/watch?v=URvWSsAgtJE)


## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.2-slides")
```

```{=html}
<!--## <i class="fas fa-laptop-code"></i> Assignments

### Preliminary Math Survey Due Next Class

Please take the [preliminary survey](/assignment/00-preliminary-survey) on your math background by next class. This will help us all have a productive semester together.-->
```
