---
title: "1.3 — The Ricardian View of Trade: Comparative Advantage — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Date

*Tuesday, January 31, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we begin looking at David Ricardo's view of trade, focusing on comparative advantage. We will derive a very simple model of comparative advantage and practice working with it. Next class, we will dive into a more formal Ricardian model that will ultimately be extended into the modern models of international trade over the coming weeks.

## <i class="fas fa-dumbbell"></i> Practice

Today you will be working on practice problems on simple comparative advantage:

- [<i class="fas fa-file-pdf"></i> Practice Problems](../files/practice/1.3-practice-pdf.pdf)
- [<i class="fas fa-file-pdf"></i> Answer Key](../files/practice/1.3-practice-answers-pdf.pdf)

## <i class="fas fa-book-reader fa-lg"></i> Readings

- Chapter 2.1 in Feenstra & Taylor

Feenstra and Taylor dive right into Ricardo's famous one-factor model starting in section 2.2. I am instead easing you into the idea of comparative advantage that doesn't focus on labor productivity and marginal products, but just some simple numbers. So you can read the first part of Chapter 2 just for a basic overview if you wish.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.3-slides")
```

```{=html}
<!--## <i class="fas fa-laptop-code"></i> Assignments

### Preliminary Math Survey Due Next Class

Please take the [preliminary survey](/assignment/00-preliminary-survey) on your math background by next class. This will help us all have a productive semester together.-->
```
