---
title: "1.5 — The Ricardian View of Trade: Reflections on Comparative Advantage — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Date

*Tuesday, February 7, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today, we go over the Ricardian model practice from last class, and since we will not have enough time to start the next model (I would rather not split it across classes if I can help it), we will reflect on the nature of comparative advantage and its relationship to the division of labor, and international trade.

## <i class="fas fa-book-reader fa-lg"></i> Readings

There are no required or unique readings for today. Though as I mentioned in class, you might find [How the Dismal Science Got its Name](https://www.econlib.org/library/Columns/LevyPeartdismal.html) insightful and surprising. It is a 6 part series of blog posts based on research by David Levy & Sandra Peart.

- [Part I](https://www.econlib.org/library/Columns/LevyPeartdismal.html)
- [Part II](https://www.econlib.org/library/Columns/LevyPeartdismal2.html)
- [Part III](https://www.econlib.org/library/Columns/LevyPeartdismal3.html)
- [Part IV](https://www.econlib.org/library/Columns/LevyPeartdismal4.html)
- [Part V](https://www.econlib.org/library/Columns/LevyPeartdismal5.html)
- [Part VI](https://www.econlib.org/library/Columns/LevyPeartdismal6.html)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.5-slides")
```

```{=html}
<!--## <i class="fas fa-laptop-code"></i> Assignments

### Preliminary Math Survey Due Next Class

Please take the [preliminary survey](/assignment/00-preliminary-survey) on your math background by next class. This will help us all have a productive semester together.-->
```
