---
title: "1.7 — The Specific Factors Model — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Date

*Thursday, February 16, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today, we cover the specific factors model.

## <i class="fas fa-book-reader fa-lg"></i> Readings

Feenstra and Taylor do a good job discussing the specific factors model.

- Chapter 3 in Feenstra & Taylor

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.8-slides")
```

```{=html}
<!--## <i class="fas fa-laptop-code"></i> Assignments

### Preliminary Math Survey Due Next Class

Please take the [preliminary survey](/assignment/00-preliminary-survey) on your math background by next class. This will help us all have a productive semester together.-->
```
