---
title: "2.1 — Tariffs — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Date

*Thursday, March 8, 2023*
:::

::: callout-warning
# Upcoming Assignment
The [Midterm Exam](/assignment/midterm) is due by 11:59 PM Friday March 10 by email.
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today, we begin trade policies by looking at the ubiquitous form of trade policy: tariffs (taxes) on imports. We might not finish today, and wrap up a few more things about tariffs next class, where we examine other types of trade barriers like quotas and subsidies.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- Chapter 8.2-8.4 in Feenstra & Taylor

We will cover the material in 8.1 (trade agreements) at the end of the course, and 8.5 (quotas) next class.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("2.1-slides")
```

## <i class="fas fa-laptop-code"></i> Assignments

### Midterm Exam Due Friday March 10

The [Midterm Exam](/assignment/midterm) is due by 11:59 PM Friday March 10 by email.
