---
title: "2.2 — Quotas, Subsidies, and Non-Tariff Barriers — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Date

*Thursday, March 8, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today, we examine other types of trade barriers like quotas and subsidies.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- Chapter 8.5-8.6 in Feenstra & Taylor

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("2.2-slides")
```