---
title: "2.5 — An Intellectual History of Trade III: Better Protectionist Arguments — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Date

*Tuesday, April 4, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today, we continue looking at an intellectual history of trade, discussing new protectionist arguments that emerged since the decline of mercantilism.

It has probably been drilled in your head by economists that tariffs are bad (and I am happy to be only the latest professor doing it!). The overwhelming majority of economists agree that tariffs, in the practical, political, real world are not a net gain. The most sober-minded and non-ideological of them recognize that there are theoretical conditions where tariffs can be good for an economy - though identifying those conditions in the complexity of the real world is difficult.

I want to give the protectionists their due, and to that end, in this discussion (and the following one, on industrial policy), I want to push you all to confront two possibilities and grapple with them:

1. There are conditions, however theoretical, where tariffs (or other trade restrictions, and not free trade) might be good.
2. Historically, the wealthy developed countries (e.g. Britain and the United States), while touting the ideology of free trade, were themselves quite protectionist for a long time. 

The opponents of free trade have latched onto these facts historically (and today), and have historically made arguments that are a variation of: the U.S./U.K. became the wealthiest countries in the world BECAUSE of their tariffs, particularly in the 19th century. This has been a significant intellectual argument among certain elements of the economic nationalist right in this country, and this was well before President Trump arrived on the scene! Later, I will trace this argument from Alexander Hamilton in the 1790s through Friedrich List and Henry Clay in the 1830s to Pat Buchanan in the 1990s. I want us to grapple with this argument, and we will do so again in our following discussion on industrial policy (they are all related).

## <i class="fas fa-book-reader fa-lg"></i> Readings

### Required Readings

- [<i class="fas fa-external-link-square-alt"></i> Nye, 2003, "The Myth of Free Trade Britain"](https://www.econlib.org/library/Columns/y2003/Nyefreetrade.html)
- [<i class="fas fa-podcast"></i> *Econtalk*: Nye on Wine, War and Trade"](https://www.econtalk.org/nye-on-wine-war-and-trade/)
- [<i class="fas fa-file-pdf"></i> Lind, 2003, "Free Trade Fallacy"](/readings/Lind-2003.pdf)
- [<i class="fas fa-file-pdf"></i> Irwin, 2002, "Interpreting the Tariff–Growth Correlation of the Late 19th Century"](/readings/Irwin-2002.pdf)

### Optional Readings

Much of my intellectual history series of lectures comes from Chapters 5-9 of Douglas Irwin’s^[One of the best economic historians on trade.] excellent book:

- [<i class="fas fa-book"></i> Irwin, 1996, *Against the Tide: An Intellectual History of Free Trade*](https://www.amazon.com/Against-Tide-Intellectual-History-Trade/dp/0691058962)

### Questions To Help Your Reading

- What were the available means for governments to raise revenues over the last few centuries? How and why did they choose certain policies? (Nye and Irwin papers)
- What is the "American system" described by Lind? How was it supposed to work? 
- According to Nye, how were England and France different in their approach to mercantilism? Hint: what were the differences in their internal vs. external markets. 
- Also interesting in Nye (more the podcast than the paper), why is/was alcohol such an important good? 

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("2.5-slides")
```