---
title: "2.7 — The Political Economy of Trade Policy — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Date

*Thursday, April 13, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today, we look at the politics of trade policy, examining how different groups in a modern democracy have different objectives and face different incentives, culminating in trade policy that concentrates benefits and disperses costs. The tools we use are standard in [“public choice”](https://en.wikipedia.org/wiki/Public_choice) analysis — the economic analysis of politics. You can learn more in [my public economics course](https://publics20.classes.ryansafner.com).

### <i class="fas fa-book-reader fa-lg"></i> Readings

### Required Readings:

- [<i class="fas fa-file-pdf"></i> Caplan, 2007, "The Myth of the Rational Voter"](/files/readings/Caplan-2007.pdf)
- [<i class="fas fa-file-pdf"></i> Somin, 2014, "When Ignorance Isn't Bliss"](/files/readings/Somin-2014.pdf)
- [<i class="fas fa-file-pdf"></i> Buchanan, 1984, "Politics Without Romance"](/files/readings/Buchanan-1984.pdf)
- [<i class="fas fa-file.pdf"></i> Mitchell, 2013, "The Pathology of Privilege: The Economic Consequences of Government Favoritism"](/files/readings/Mitchell-2013.pdf)

### Optional Readings

- [<i class="fas fa-file-pdf"></i> Grossman and Helpman, 1994, "Protection for Sale"](/files/readings/Grossman-Helpman-1994.pdf)
- [<i class="fas fa-file-pdf"></i> Baldwin and Magee, 2000, "Is trade policy for sale? Congressional voting on recent t?"](/files/readings/Baldwin-Magee-2000.pdf)

### Helpful:

- [<i class="fab fa-wikipedia-w"></i> Public Choice](https://en.wikipedia.org/wiki/Public_choice)
- [<i class="fab fa-wikipedia-w"></i> The Logic of Collective Action](https://en.wikipedia.org/wiki/The_Logic_of_Collective_Action)
- [<i class="fab fa-wikipedia-w"></i> Government failure](https://en.wikipedia.org/wiki/Government_failure)
- [<i class="fab fa-wikipedia-w"></i> Rational ignorance](https://en.wikipedia.org/wiki/Rational_ignorance)
- [<i class="fab fa-wikipedia-w"></i> Rent seeking](https://en.wikipedia.org/wiki/Rent_seeking)
- [<i class="fab fa-wikipedia-w"></i> Regulatory capture](https://en.wikipedia.org/wiki/Regulatory_capture)

### Questions To Help Your Reading

- Why is good governance a public good, subject to a free rider problem?
- What is “rational ignorance”?
- What is Caplan’s idea of “rational irrationality” and how is it different from “rational ignorance?”
- According to Caplan, what are the four systematic biases?
- Why do people vote or not vote? How do people that do vote choose how they vote?
- What is the principle of “concentrated benefits and dispersed costs?”

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("2.7-slides")
```