---
title: "3.2 — Globalization and its Discontents — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Date

*Thursday, April 27, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

### <i class="fas fa-book-reader fa-lg"></i> Readings

- [<i class="fas fa-external-link-square-alt"></i> Tarko and Safner, 2022, "International regulatory diversity over50 years: political entrepreneurship withinfiscal constraints "](https://link.springer.com/epdf/10.1007/s11127-022-01011-2?sharing_token=0PMazGn6dpbw8WWmSOLsYPe4RwlQNchNByi7wbcMAY6sM4iEqIYI0dMv-E9PvzDaltt8-HxXVpJtBBw8N3WxXoQuXY6rFWHLNbeEmS-MZlq11VW8XneB__PPi3kRJ_nLR07J2ZIZcwmQkHNQ9JN-0-v3DX27mJDdJ8pw8OE_P-U%3D)

- [<i class="fas fa-external-link-square-alt"></i> Bremmer, 2017, "In Defense of the Trump Voter "](https://www.linkedin.com/pulse/defense-trump-voter-ian-bremmer/)
- [<i class="fas fa-external-link-square-alt"></i> Hall, 2017, "Six Theses about Contemporary Populism"](https://gem.cid.harvard.edu/files/gem2016/files/hall_blyth_berglof_gem17.pdf)
- [<i class="fas fa-external-link-square-alt"></i> WSJ, Nov. 10, 2020, "How Politics Has Pulled the Country in Different Directions"](https://www.wsj.com/graphics/polarized-presidential-elections/?mod=hp_lead_pos9)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("3.2-slides")
```